### 2019-12-11 ms/Magalhaes_Schwilk_litter_flam_moisture.tex
Work on edits to introduction
in response to reviews

%% Revision 2019-12-10
\documentclass[letterpaper,12pt]{article}
\usepackage{jecol}

\title{Moisture effect on leaf litter alters non-additive flammability in mixed conifer forest}

\running{Leaf litter flammability}

\author{Rita Quinones de Magalhaes$^{1,2}$ and Dylan W. Schwilk$^1$}

\affiliations{
    \item Texas Tech University, Lubbock, Texas 79411, USA.
    \item Current address: Rochester Institute of Technology, Rochester, NY 14623, USA.
}

\nwords{7600}
\ntables{4}
\nfig{8}
\nref{39}

\corr{\url{dylan.schwilk@ttu.edu}}

\begin{document}
\maketitle

\begin{abstract}
  \noindent \begin{enumerate}

      \item Moisture content is a strong determinant of forest fuel
    flammability, however, little is known about the mechanisms by which
    species-specific litter traits influence litter moisture absorption and
    retention. Characterizing how species’ leaf litter retain moisture through
    time and how fuel from multiple species interacts are critical to the
    assessment of fuel’s availability to burn and key steps to adding species
    or taxon specific information to fire behavior prediction.

      \item To fill these gaps in knowledge, determined
    moisture absorption and retention characteristics of the litter from
    multiple tree species, investigated what litter traits influenced these
    processes, and tested for nonadditive moisture dynamics and nonadditive
    flammability in multi-species mixtures. The litter from eight tree species of a
    mixed-conifer forest was saturated and allowed to dry to determine moisture
    absorption capacity and drying rates. Burn trials were performed in moist
    litter beds of single species and of mixtures to establish flammability
    response to the dry-down process.

      \item Moisture dynamics varied across species with higher specific leaf
    area being associated with higher maximum absorption and lower litter bulk
    density being associated with faster drying rates. We examined two axes of
    flammability variation and found that these differing moisture dynamics
    result in time since wetting having strongly differing effects on
    flammability across the major litter types typified by different tree taxa.
    Litter mixtures exhibited consistent nonadditivity in flame spread rate and
    in fuel consumption. This nonadditivity became more positive with fuel
    drying.

      \item \emph{Synthesis:} Taxon specific patterns of moisture absorption
    and retention can change flammability rankings under moist compared with dry fuels.
    Litter mixtures generally burn with fire behavior more similar to that of
    the most flammable constituent species and this effect increases as fuels
    dry. [TODO more]

  \end{enumerate}
\end{abstract}

\noindent \textbf{Keywords:} California, fire, flame spread rate, flammability, fuel consumption, fuel moisture, leaf litter, nonadditivity

\newpage

\section*{Introduction}

Fuel moisture content is an important, and often driving, determinant of
surface fire behavior \citep{Rothermel-1972}. Moisture acts as a heat-sink
resulting in less energy available for propagation, and the resulting water
vapor will both cool the flaming front and dilute flammable gases
\citep{Albini-1976, Shafizadeh-1977}. Environmental conditions such as
temperature, relative humidity, and wind drive the moisture content of the fuel
available to burn \citep{Kreye+Varner+etal-2018}. Variation in moisture of
living fuels has strong effects on flammability, but moisture of dead excised
litter fuels can also vary. In particular, fine dead fuels respond very readily
to environmental changes \citep{Nelson-2001}. Fuel moisture influences
flammability and fire behavior through effects on ignition, spread rate, flame
height, fuel consumption, and heat release, due to the high specific heat
content of water \citep{Rothermel-1972, Nelson-2001}. Moisture of litter fuels
varies with climate, with seasons, and with short term weather within a season.
Therefore, the rate of drying and flammability as a function of moisture
content are key parameters that influence fuel flammability throughout the fire
season \citep{Nelson-2001}.

Flammability is a multi-dimensional characteristic of fuels
\citep{Schwilk-2015, Pausas+Keeley+etal-2017} with different traits influencing
different components of fire behavior. In many forests that experience surface
fire, litter properties influence fire behavior
\citep{Ganteaume+Marielle+etal-2011, Schwilk+Caprio-2011,
  Varner+Kane+etal-2015}. In litter driven fire, the two major axes of
variation are described by measurements related to flame spread rate and
measurements related to heat release or heat transfer
\citep{Magalhaes+Schwilk-2012, Prior+Murphy+etal-2018}. These different
components of fire behavior have differing ecological effects with spread rate
being associated with rates of heat release and total burned area and with heat
transfer to soils or tissues predicting soil impacts of fire and the mortality
of better insulated plant tissues. A shortcoming in our current knowledge is
our integration of specific species and plant traits into fuel hazard
assessment, fire modeling, and fire effects prediction. Developing quantitative
yet robust and general ways to integrate actual plant communities into fuel
assessment will require understanding the magnitude of flammability variability
across species in space and time and parameterizing based on plant
characteristics. [TODO: get opinion on final sentences]

Species differ in their litter flammability \cite{Fonda+Belanger+etal-1998,
  Fonda-2001, Scarf+Westoby-2006, Kane+Varner+etal-2008,
  Cornwell+Elvira+etal-2015}. Litter packing, measured as litter-bed density or
packing ratio, is an important control ignition and spread in models of surface
fire spread \citep{Weber-1991, Balbi+Santoni+etal-1999, Morvan+Larini-2001}.
Multiple studies have demonstrated that litter bed density controls the
flammability of dry litter fuels in seasonally dry climates where higher litter
density reduces flame spread rate. Across these studies, there is increasing
consensus that larger less densely packed litter beds increase oxygen
availability and burn with faster spread rates \citep{Scarf+Westoby-2006,
  Schwilk+Caprio-2011, Magalhaes+Schwilk-2012, Cornwell+Elvira+etal-2015,
  Varner+Kane+etal-2015, Kreye+Varner+etal-2018}. The litter packing is driven
by litter particle size and shape which is in turn influenced by leaf size, how
leaves or branches are abscised, and subsequent decomposition
\citep{Fonda+Belanger+etal-1998, Stephens+Finney+etal-2004,
  Scarff+Westoby-2006, Kane+Varner+etal-2008, Cornwell+Elvira+etal-2015,
  Zhao+Cornwell+etal-2016, Grootemaat+Wright+etal-2017,
  Prior+Murphy+etal-2018}.

These species-specific effects are increasingly well established. However, how
moisture dynamics modulate species specific flammability is poorly known and
could be driven by by leaf traits, emergent litter traits, or by both. Species
specific effects on litter flammability are recorded for the forests of the
southeastern US \citep{Nowacki+Abrams-2008} where species-specific litter
drying rates play a role in flammability feedbacks
\citep{Kreye+Varner+etal-2013}. Fire behavior consequences of individual
species traits have been recorded for western US forests as well: in mixed
conifer forest in California, where the fire regime is characterized by surface
fires fueled by plant litter, overstory tree composition drives variation in
local fire behavior \citep{Schwilk+Caprio-2011} as a result of differential
litter flammability across tree species and as a result of synergistic
interactions in multi-species litter mixtures \citep{Magalhaes+Schwilk-2012}.
These flammability feedbacks may help explain historical fire frequency
patterns \citep{Schwilk+Caprio-2011}, but current estimates of litter
flammability in mixed conifer forests are based largely on burning dry litter.

Fire scientists often categorize fuel by size classes that correspond to
moisture exchange rates, although such simplified categories have long been
recognized as not sufficient to describe species-specific effects on moisture
dynamics \citep{Anderson-1985}. Fuel moisture dynamics of leaf litter are
dependent upon forest structure (shading and rain throughfall effects), litter
physical structure, and chemical composition of the fuel-beds
\citep{Nelson+Hiers-2008, Matthews-2014, Kreye_Hiers_etal-2018}. A variety of
models describing moisture dynamics of fine fuels are used by fire scientists
\citep{Viney-1991, Nelson-2000, Catchpole+Catchpole+etal-2001}. These models
usually aim to describe equilibrium conditions and at least include temperature
and relative humidity although some include precipitation and solar radiation
effects as well. Explicit models of fire behavior as a function of time since
wetting could improve fire prediction, however. Drying rates following snowmelt
or following rainfall during the fire season will influence the length of the
fire season, flammability over time, and the probability of sustained ignition,
higher spread rates and fuel consumption when ignition occurs. That species can
differ in litter drying rates has been demonstrated for at least one forest
system \citep{Kreye+Varner+etal-2013}. Such differential moisture content
across time may contribute to the often high spatial heterogeneity of surface
fires \citep{Knapp_Schwilk_etal-2006, Kreye_Hiers_etal-2018}.

Fuel moisture content through time is controlled by two processes, absorption
and retention \citep{Kreye+Kobziar+etal-2013}. Absorptive capacity represents
the potential maximum fuel moisture and is evaluated by the moisture content of
litter when saturated. Differences in moisture content over time since
saturation can indicate different abilities to retain moisture. Litter particle
geometry may have direct effects on maximum moisture absorption through
area-mass relationships and we also expect effects of leaf morphology,
especially leaf cuticle thickness \citep{Van_Wagner-1969}. Like absorption,
moisture retention should also be influenced by leaf and litter bed traits:
surface-area-to-volume ratio and litter bulk density will influence air flow
through the litter bed. Litter packing therefore effects fire behavior in two
ways: directly by its effect on oxygen availability for combustion, and
indirectly by its effect on dry-down rate. Both these pathways are affected by
leaf particle traits, especially leaf size \citep{Scarff+Westoby-2006}.
Fuel-beds composed of larger leaves with lower moisture retainability should
favor drying, thus increasing the probability of ignition and fire spread.

In mixed species communities, species traits may interact such that the
flammability of multi-species mixtures is not the average of the constituent
species. Such nonadditive effects are often synergistic in dry litter fuels
where the flammability of a mixture is driven by the most flammable species
\citep{VanAltena+Logtestjin+etal-2012, Magalhaes+Schwilk-2012}, but negative
nonadditivity has also been recorded for litter mixtures
\citep{Blauw+Wensink+etal-2015, Zhao+Cornwell+etal-2016}. It is possible that
similar nonadditive effects may influence litter moisture content directly and
flammability of moist fuels. There is some evidence for nonadditive
flammability in moist fuels with one study finding increasing nonadditivity
with moisture \citep{Blauw+Wensink+etal-2015}.

[TODO: reseearch question here]

We investigated species specific litter moisture dynamics across eight tree
species in a temperate fire-prone forest in California. We determined the
maximum water retention and rate of moisture loss and examined how these vary
in relation to two leaf and litter bed traits [CHECK AND FIX]. We then examined
the influence of moisture on the flammability across these species.
Experimental burn trials of leaf litter at various moisture levels determined
how time since wetting influenced two axes of flammability characterized by
flame spread rate and total fuel consumed. Previous work described the
flammability of these species in monocultures and in mixtures for oven-dry
litter \citep{Magalhaes+Schwilk-2012} and found consistent positive
nonadditivity of flammability where fire behavior of mixtures was driven by the
most flammable species. Here, we examined if this positive nonadditivity was
maintained in mixtures of varying levels of fuel moisture. The burn trials also
performed with multi-species litter mixtures to evaluate if there are
nonadditive effects as observed in oven-dried litter. We tested if the
nonadditive effects on reported in previous studies occur under [TODO CHECK]
ecologically-relevant moisture levels, and whether fuel moisture effects
strengthen or moderate such nonadditivity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Materials and Methods}

\subsection*{Site description and species selection}

Field sites for this study were located in Sequoia and Kings Canyon National
Parks, California, USA (N$36^\circ 36'$, W$118^\circ 42'$) between 1600 and
2400\,m elevation in mixed-conifer forest. We chose eight dominant tree species
that are representative of this type of forest: \emph{Pinus jeffreyi} Grev. \&
Balf., \emph{Pinus lambertiana} Dougl., \emph{Pinus ponderosa} Dougl. ex Laws.,
\emph{Abies concolor} (Gord. \& Glend.) Lindl. ex Hildebr., \emph{Abies
  magnifica} A. Murr., \emph{Calocedrus decurrens} (Torr.) Florin,
\emph{Quercus kelloggii} Newb., and \emph{Sequoiadendron giganteum} (Lindl.) J.
Buchholz. In this study, we consider litter the top layer of leaves and small
twigs less than 0.625\,cm diameter (= 1-hour fuel) that have fallen in the
previous year (mostly undecomposed or lightly decomposed). Leaf litter was
collected in the summer (mid-June to mid-July) in 2012 from 21 separate sites
across the study area. We collected from a minimum of four populations per
species with each population 9--32\,km distant from its neighbors. An exception
to this was \emph{A. magnifica} which is restricted to the higher elevations
and could only be collected from one area due to designated wilderness
boundaries and logistical constraints. Within that area, \emph{A. magnifica}
litter was collected from over 10 different individuals. For the remaining
species, the collection involved 2--4 individual trees at each site which were
separated from one another by at least 10\,m. The litter was collected 2\,m
away from the trunk to obtain a more uniform sample of leaf and fine branchlet
litter, because bark and heavier branches tend to fall closer to the tree
trunk. The top 2.5-5\,cm of litter were collected by raking and shoveling into
large bags. All litter was air dried to $< 5$\% fuel moisture by dry weight,
then weighed and divided into moisture trials. The collected litter was dried
in large batches, then stored in sealed plastic bags inside cardboard boxes in
a lab storage space. We minimized handling and moving the bagged litter to
avoid mechanical damage and changes to the particle size distributions. The
moisture manipulations and subsequent burning trials were conducted in
April--August 2015.

\subsection*{Moisture absorption and desorption}

We produced dry-down curves to determine maximum water retainability and the
rate of moisture content loss for each species and for four distinct
three-species mixtures. These curves allowed us to determine the rate at which
a species loses moisture, approximated as exponential decay, and to determine
maximum water content following saturation and gravity draining. Samples from
the eight species (six replicates per species) and four mixtures (5 replicates
per mixture type) were placed in baskets approximately 45 x 45\,cm by 15\,cm
tall constructed of aluminum mesh with 0.52\,cm square openings. Samples were
450\,g dry weight litter per trial and litter depth was standardized to 10\,cm
by changing the horizontal dimensions of the basket. Weight was measured by
using a balance sensitive to 0.1\,g (model XS16001L, Mettler Toledo, Columbus,
OH). Samples were immersed in plastic storage bins filled with water for 24
hours to fully saturate them. After saturation, the baskets were removed from
the water, allowed to drain for 3 minutes, and samples weighed to measure
maximum moisture content. The samples were then allowed to dry in a controlled
environment kept at $21^\circ$C temperature and 30\% relative humidity,
assessed via an iButton sensor (Maxim Integrated) placed over the samples.
Litter baskets were arranged in a completely randomized design. During the
dry-down process three subsamples of 5--10\,g were taken at 24 hour intervals
and weighed. These subsamples were then oven dried for 24 hours at $100^\circ$C
and re-weighed to assess moisture content. This was repeated until the final
sample from a basket was less than 10\% moisture.

With only eight species in this study, our power to test leaf trait effects on
the observed moisture dynamics was low, so we decided \emph{a priori} to test
the effect of one leaf trait likely to influence moisture absorption (specific
leaf area), and one litter trait known to directly affect flammability and
likely to influence drying rate and litter packing density (litter bulk
density). [TODO check] Specific leaf area was measured on freshly fallen litter
by sampling 5 individuals per species. This measurement included petiole and
for the scale-leaved species included the full excised branchlet. Therefore,
this is not directly comparable to typical specific leaf area measurements on
fresh leaves. Litter bulk density was obtained in this experiment for each
basket by dividing the dry mass (450\,g) by the estimated litter volume (basket
width x length x litter depth).

\subsection*{Flammability assessment}

We performed flammability tests across multiple moisture levels on monospecific
litter beds and on mixed litter beds with each species or mixture replicated 5
times at each moisture level. We used the initial dry-down experiment to guide
our timing so as to attempt to burn samples over a range of moisture contents.
We aimed for moisture contents from 10\% to 100\% on a dry mass basis as
preliminary trials on damp fuels suggested that across these fuels ignition
will not occur above 100\% fuel moisture. To expand our data on the dry end of
this gradient, we supplemented these data with data from an earlier experiment
on oven-dried litter \citep{Magalhaes+Schwilk-2012}. This supplemental data
added five replicates per species at an oven dried moisture content of $<$ 5\%.
We assigned these replicates a time since wetting value of 144 hours equivalent
to the longest time since drying for the wetted samples. Flammability was
assessed using a 150\,cm long burn table, 15\,cm wide and 15\,cm tall, in which
leaf litter was placed and the table gently shaken to allow settling. Burning
trials all used 450\,g (dry weight) of litter. For three-species mixtures, this
was 150\,g per species. [TODO give depths] This design allows fire to reach
constant flame spread rates \citep{Magalhaes+Schwilk-2012}. Two graduated metal
rulers equally spaced along the apparatus allowed visual determination of
maximum flame height. Litter was ignited with a propane torch; maximum
adiabatic flame temperature of 1899$^\circ$C,
\url{http://www.benzomatic.com/products/fuel.aspx}) at one end of the apparatus
and allowed to burn until extinction. If a sample failed to ignite after 30
seconds, no further attempt was made and if three samples of a species at a
particular moisture level failed to ignite, no further attempts were made on
that species/moisture combination and the replicate was allowed to continue
drying until the next scheduled checkpoint. A timer was used to record total
time to ignition and duration of flaming combustion. The flammability trials
followed the methodology detailed in \citet{Magalhaes+Schwilk-2012}.

During flammability trials, we recorded time to ignition (s), duration of
flaming combustion (s), flame spread rate (mm/s), calculated by dividing the
length of the burned surface by the time it takes the flaming front to reach
the end of the apparatus, maximal flame height (mm), and fuel consumption (\%).
Some trials resulted only in smoldering combustion; for those trials, a time to
ignition and fuel consumption were recorded, but flame spread rate and flame
height were recorded as zero. Our thermocouple system failed or data was lost
for two days of burning trials, therefore we have omitted those data from this
analysis. Our past work \citep{Magalhaes+Schwilk-2012} demonstrated that flame
spread rate captures one important axis of flammability variation
\citep{Schwilk-2015, Pausas+Keeley+etal-2017, Prior+Murphy+etal02018} and
percent fuel consumed captures a portion of the second major axis of
flammability variation. Our analysis, therefore, focused on these two
relatively orthogonal measures of flammability to avoid examining multiple
redundant dependent variables.

Burn trials occurred in a cement structure used to simulate house fires at the
Fire Department of the City of Lubbock. This structure minimized wind and
helped regulate temperature and relative humidity. Temperature and relative
humidity were measured before every trial using a Kestrel 3000
(Nielsen-Kellerman, Boothwyn, PA). All trials were conducted on clear days
between 10 am and 3 pm.

\subsection*{Statistical analyses}

Dry-down curves were created based on the drying experiment, fitted with an
exponential decay curve to each species.:

\begin{equation}
m = m_{max} e^{-\lambda t}
\end{equation}

Where $m$ is moisture content at time $t$, and $\lambda$ is the exponential
decay coefficient. To estimate these coefficients we fit a linear mixed-effects
model by first taking the natural logarithm of moisture content, with time
since wetting and species as fixed effects nested within replicate basket. We
extracted the coefficients of the fitted model to describe maximum water
retainability ($m_{max}$; g of water per g of dry mass) and desiccation rate
($\lambda$; hour$^{-1}$) for each species. All analyses were conducted in R
3.6.1 \citep{RCoreTeam-2019}. We fit models with R packages `lme4'
\citep{Bates_Machler_etal-2015} and `afex' \citep{Singmann_Bolker_etal-2017}.
Approximate degrees of freedom and p-values were calculated by the
Kenward-Roger approximation \citep{Kenward_Roger-1997} which is recommended by
Luke \citeyear{Luke-2017} as producing acceptable Type I error rates. We tested
if one litter particle trait (specific leaf area) and one litter bed trait
(litter bulk density) influenced moisture dynamics by fitting linear models
with species means as observations. For each response variable we fit a single
linear model with two predictors and an interaction term.

Our past work \citep{Magalhaes+Schwilk-2012} demonstrated that dry litter
flammability across these eight species was driven by leaf traits and litter
packing. These previously reported flammability trials demonstrated that
species within a genus tended to behave similarly and the two scale-leaved
genera in the Cupressaceae, \emph{Calocedrus} and \emph{Sequoiadendron}, also
behaved similarly to one another. Therefore, to preserve power we used
taxonomic groups (\emph{Abies}, Cupressaceae, \emph{Quercus}, and \emph{Pinus})
as our flammability functional groups and treated these as fixed effects in our
models. The effect of fuel moisture and time since wetting on flammability
parameters were modeled using linear models with fuel moisture or time since
wetting and taxon as fixed effects and average vapor pressure deficit in the
burning room during that trial as a nuisance covariate. We used linear mixed
models to predict spread rate, but fuel consumption tends to exhibit threshold
effects with few intermediate values. Therefore, we used mixed generalized
linear models with a beta regression link with the `glmmADMB' package
\citep{Skaug_Fournier_etal-2016} to model percent fuel consumed as a function
of moisture content and of time since wetting.

We investigated possible nonadditive effects on moisture and on flammability by
testing if moisture content and flammability of mixtures were predicted by the
average of the constituent species. To choose mixtures, we first grouped the
species according to their maximum water retention and drying rate according to
estimated marginal means by species using the R package `emmeans'
\citep{Lenth-2019}. This resulted in four groups of species and we then
selected one representative species from each litter type group and produced
all four possible three-species litter mixtures using those four species. For
each mixture and time since wetting, we predicted moisture content based on the
mean content of the three constituent species at that time since wetting. We
predicted flammability parameters based on the average flammability measures
from individual species trials at each time since wetting and then calculated a
predicted spread rate and fuel consumption as the average of these three values
for the mixture. The expectation from the null model was that the difference
between observed and predicted values of flammability will be zero. We tested
if there were significant departures from the null with a mixed effects linear
model. For flammability parameters, we fit linear mixed effect models with the
observed-predicted flammability value as the dependent variable and used Wald
tests to test if the centered intercept was different than zero (a measure of
overall nonadditivity) and if the slope with time since wetting was different
from zero (indicating nonadditivity changing with drying).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}

\subsection*{Species specific moisture dynamics and leaf traits}

There were marked differences in the extent of moisture species retain
initially following wetting and draining by gravity (all $P < 0.0001$, Fig.
\ref{fig:drydown} and Table \ref{tab:drydown}). \emph{Q. kelloggii}, the only
broadleaf, is the species that retained the most moisture initially.
\emph{Quercus}, like the pines, had a rapid drying rate (Fig. \ref{fig:drydown}
and SI Fig. 1). Calculating estimated marginal means following the linear
mixed-effects model discriminated the species into three different maximum
water retention groups: 1) \emph{Q. kelloggii}; 2) \emph{P. lambertiana} and
\emph{C. decurrens}; and 3) \emph{A. magnifica}, \emph{S. giganteum}, \emph{A.
  concolor}, \emph{P. jeffreyi} and \emph{P. ponderosa} (SI Fig. 1). There are
also differences in drying rates across the eight species and estimated mixed
linear model slopes discriminated species into three groups from lowest to most
rapid: 1) \emph{C. decurrens}, 2) \emph{Abies spp} and \emph{S. giganteum}, and
3) \emph{Q. kelloggii} and \emph{Pinus spp} (SI Fig. 2).

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig1_drydown-curves.pdf}
\caption{Dry down curves for eight litter types. Lines show best-fit
  exponential decay curves fit to moisture content on a dry mass basis. Litter
  samples we saturated by immersion for 24 hours and then allowed to drain by
  gravity for 3 min before initial weighing. Colors indicate genera and there
  were significant differences in both maximum water retention and in rate of
  drying (based linear model on log of water content). A small amount of
  horizontal jitter (0.5 hr) was added to this plot to aid in visualizing
  because, although collections were staggered, there were sometimes multiple
  samples collected at a single time point.}
 \label{fig:drydown}
\end{figure}

\begin{table}[h]
  \caption{Linear mixed model results for moisture content as a function of
    time since wetting and species. Approximate degrees of freedom (df), pseudo F
    statistics and p-values were calculated by the Kenward-Roger approximation.
    \citep{Kenward_Roger-1997}. Model coefficients reported in SI Table 1.}
  \label{tab:drydown}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{llll}
  \toprule
Effect & df & $F$ & $P$ \\ 
  \midrule
  hour & 1, 183.73 & 1599.59 & $<$.0001 \\ 
  species & 7, 89.73 & 36.15 & $<$.0001 \\ 
  hour:species & 7, 184.15 & 23.06 & $<$.0001 \\ 
   \bottomrule
\end{tabular}
\end{table}

Across these eight species specific leaf area had a positive association with maximum
water retention (SI Fig 4, linear model $P$ = 0.003) and litter
bulk density had a negative effect on desiccation rate (SI Fig 5,
linear model $P$ = 0.029). ANOVA results are shown in Tables
\ref{tab:mc_di_anova}.

\begin{table}[h]
  \caption{Linear model results for maximum water retention (\% moisture
    content on dry mass basis) and for estimated drying rate (hr$^{-1}$) as
    functions of specific leaf area and litter bulk density across eight
    species. ANOVA tables shown with degrees of freedom, sum of squares, mean
    squares, $F$, and $P$ for each effect. Model coefficients in SI tables 2 and 3}
  \label{tab:mc_di_anova}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.

\begin{tabular}{ccrrrrr}
  \toprule
Ind. variable & Effect & df & Sum of Squares & Mean squares & $F$ & $P$ \\ 
  \midrule
Moisture content & SLA & 1 & 250834.5 & 250834.51 & 130.13 & 0.000 \\ 
  & density & 1 & 538.5 & 538.54 & 0.28 & 0.625 \\ 
  & SLA:density & 1 & 5135.2 & 5135.24 & 2.66 & 0.178 \\ 
  & Residuals & 4 & 7710.4 & 1927.61 &  &  \\ 
  \midrule

Drying rate & density & 1 & $3.92 \times 10^{-4}$ & $3.92 \times 10^{-4}$ & 11.01 & 0.029 \\ 
 &  SLA & 1 & $1.88 \times 10^{-5}$ & $1.88 \times 10^{-5}$ & 0.53 & 0.507 \\ 
 &  density:SLA & 1 & $6.89 \times 10^{-6}$ & $6.89 \times 10^{-6}$ & 0.19 & 0.683 \\ 
 &  Residuals & 4 & $1.42 \times 10^{-4}$ & $3.56 \times 10^{-5}$ &  &  \\ 
   \bottomrule

\end{tabular}
\end{table}

\subsection*{Moisture effects on flammability}

Flame spread rate declined with fuel moisture (Fig. \ref{fig:spread_moist}).
\emph{Quercus} had higher spread rates across all fuel moistures and had a
steeper decline in response to moisture than did the other groups (taxon by
time since wetting interaction $P$ = 0.022, Table \ref{tab:spreadrate_anova}).
Fuel moisture negatively influenced fuel consumption. Overall fuel consumption
and the rate of change with moisture content varied across the taxonomic groups
(Fig. \ref{fig:consume_moist}, Table \ref{tab:consume_anova}). Although the
overall ranking of groups was similar for fuel consumption and for flame spread
rate, in contrast to the pattern with spread rate, for fuel consumption,
\emph{Pinus} and \emph{Abies} has a steeper decline in flammability with fuel
moisture than did the other groups.

%% spread rate by moisture
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig2_spread_actualMC.pdf}
\caption{Flame spread rate by moisture content (model fit $F$= 61.15, df = 9,
  148, $P$ < 0.0001). Model ANOVA table in Table \ref{tab:spreadrate_anova} and
  coefficients in SI Table 4.}
  \label{fig:spread_moist}
\end{figure}

%% Table 3 
\begin{table}[h]
  \caption{Generalized linear mixed model results for two models predicting
    flame spread rate: the first model examined spread rate as a function of
    fuel moisture (mc) with taxonomic group as a fixed effect and species as a
    nested random effect. The second model predicted spread rate as a function
    of time since wetting (time) with taxonomic group as a fixed effect and
    species as a nested random effect. Approximate degrees of freedom (df),
    pseudo F statistics and p-values were calculated by the Kenward-Roger
    approximation \citep{Kenward_Roger-1997}. Estimated coefficients are in SI
    Tables 4 and 6}
  \label{tab:spreadrate_anova}

\centering
\begin{tabular}{ccrrrrr}
  \toprule
Model set & Effect & df & $F$ & & $P$ \\
  \midrule
  Effect of moisture & mc & 1, 146.20 & 21.04 & $<$.0001 \\ 
 & taxon & 3, 51.58 & 65.08 & $<$.0001 \\ 
 & vpd & 1, 148.79 & 9.62 & .002 \\ 
 & mc:taxon & 3, 147.21 & 4.83 & .003 \\

  \midrule

  Effect of time & hour & 1, 214.50 & 66.22 & $<$.0001 \\ 
 &  taxon & 3, 7.59 & 61.81 & $<$.0001 \\ 
 & vpd & 1, 215.00 & 13.49 & .0003 \\ 
 & hour:taxon & 3, 183.97 & 3.99 & .009 \\

   \bottomrule

\end{tabular}
\end{table}

%% fuel consumption by moisture:
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig3_consume_actualMC.pdf}
\caption{Percent fuel consumed by moisture content. Lines indicate best fit
  beta regression lines. Model deviance table in Table \ref{tab:consume_anova}
  and coefficients are in SI Table 5.}
  \label{fig:consume_moist}
\end{figure}

\begin{table}[h]
  \caption{Generalized linear mixed model results for two sets of models
    predicting fuel consumption. The first set modeled fuel consumption as a
    function of fuel moisture (mc) with taxonomic group as a fixed effect and
    species as a nested random effect. The second set modeled fuel consumption
    as a function of as a function of time since wetting (time) with taxonomic
    group as a fixed effect and species as a nested random effect. The analysis
    of deviance table below shows sequential chi-square tests on nested beta
    regression models: from the simplest null model within each group with only
    intercept and random effect, for a taxon effect only model, for a taxon and
    moisture content effect, and for the full model including the taxon by
    moisture or time interaction. Coefficient estimates are in SI Tables 5 and
    7.}
  \label{tab:consume_anova}
  \centering

  \begin{tabular}{llrrrr}
\toprule
Model set & model & Log Lik. & df & Deviance & $P$ \\ 
\midrule
  Effect of moisture &   \~{} 1 & 262.12 &  &  &  \\ 
   &  \~{} taxon & 268.73 & 3 & 13.23 & 0.0042 \\ 
   & \~{} taxon + mc & 289.57 & 1 & 41.68 & $<$ 0.0001 \\ 
   & \~{} taxon + mc + taxon:mc & 293.87 & 3 & 8.60 & 0.0351 \\ 
\midrule
Effect of time &  consum \~{} 1 & 140.08 &  &  &  \\ 
 & \~{} taxon & 155.84 & 4 & 31.51 & $<$ 0.001 \\ 
 & \~{} taxon + time & 209.84 & 1 & 108.01 & $<$ 0.001 \\ 
 & \~{} hour + taxon + taxon:time & 216.96 & 4 & 14.23 & 0.01 \\ 
    \bottomrule
    \end{tabular}
\end{table}

As a consequence of the relationships shown above, time since wetting had a
significant effect on spread rate ($P < 0.0001$) and the slope of this effect
varied by taxonomic group (Fig. \ref{fig:spread_time}, Table
\ref{tab:spreadrate_anova}). The pines were able to ignite and burn much sooner
after wetting than did the other taxa, none of which had measurable spread
rates until burning was attempted 72 hours after wetting (note that trials that
failed to even ignite smoldering combustion are omitted completely from the
model and figure). At 72 hours, the \emph{Abies}, \emph{Calocedrus}, and
\emph{Sequoiadendron} had dryer litter than did the \emph{Quercus}, yet the
spread rates were slow. \emph{Quercus} had the steepest response to time since
wetting (Fig. \ref{fig:spread_time}, SI Table 6)).

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig4_spread_time.pdf}
\caption{Flame spread rate by time since wetting for litter of eight species.
  Lines indicate best fit linear model fits. Model ANOVA table in Table
  \ref{tab:spreadrate_anova} and coefficients in SI Table 6. Horizontal jitter
  (up to 2 hr) was added to the plot for visualization and to avoid
  overplotting. All collections were at the times labeled on the x-axis. A lack
  of data at earlier time points for a taxon indicates no ignitions occurred at
  those time points.}
\label{fig:spread_time}
\end{figure}

Fuel consumption was highly variable across trials and intermediate values were
less common than were very low or moderately high consumption (Fig.
\ref{fig:consume_time}). All parameters significantly contributed to the
generalized linear mixed beta regression model as indicated by the nested model
comparisons in Table \ref{tab:consume_anova}. As with flame spread rate some
pines had the most rapid response to drying. However, unlike in the other
flammability analyses, here the pines did not all behave similarly: \emph{P.
  lambertiana} had lower fuel consumption than did \emph{P. ponderosa} and
\emph{P. jeffreyi}. We noticed this during trials and in model fitting we found
that placing \emph{P. lambertiana} in its own group allowed model convergence
that was otherwise impossible. This was at odds with our intent to use only
\emph{a priori} taxonomic groups, but we could not obtain model convergence
otherwise. This result should therefore be seen as preliminary because the test
does not reflect an \emph{a priori} hypothesis.

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig5_consume_time.pdf}
\caption{Percent fuel consumed by time since wetting for litter of eight
  species. Curves shown indicate best fit beta regression lines. Model ANOVA
  table in Table \ref{tab:consume_anova} and coefficients in SI Table 7.
  Horizontal jitter (up to 2 hr) was added to the plot for visualization and to
  avoid overplotting. All collections were at the times labeled on the x-axis..
  A lack of data at earlier time points for a taxon indicates no ignitions
  occurred at those time points.}
  \label{fig:consume_time}
\end{figure}

\subsection*{Nonadditivity in  litter mixtures}

Based on visual inspection of the dry down curves and the estimated marginal
means for maximum water retainability and desiccation rate (SI Figs 2--3), we
selected four species to represent the full range of litter moisture dynamics
for use in the litter mixtures and examined all four three-species combinations
of those four. The four species were \emph{Abies concolor}, \emph{Calocedrus
  decurrens}, \emph{Pinus jeffreyi} and \emph{Quercus kelloggii}.

Mixture dry down rates and maximum water retention showed some relationship to
the dry down curves for individual constituent species (see SI Fig. 6). For
example, the mixture that does not contain \emph{Q. kelloggii} (AbCaPi in SI
Fig. 6) has the lowest initial retention. Initial water retention and drying rates
differed across mixtures (linear model based on log of moisture content, SI Tables 8 and 9).

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig6_mixture_obs_vs_pred_mc.pdf}
\caption{Observed vs predicted fuel moisture content for litter mixtures.
  Predictions are based on the mean moisture content at each time since wetting
  for the three constituent species in each mixture. Codes for species in
  mixtures: Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_mc}
\end{figure}

There was evidence for nonadditive effects in drying rate. Mixtures tended to
have lower water content than predicted from the mean of the constituent
species after wetting, but had higher water content than expected after long
drying (Wald test shows slope of mixed linear model illustrated in Fig.
\ref{fig:mixture_obs_pred_mc} less than 1, $P < 0.0001$).

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig7_mixture_obs_vs_pred_spread.pdf}
\caption{Observed vs. predicted flame spread rate for litter mixtures. Filled
  circles show observed spread rate values. Open diamonds indicate predicted
  spread rates based on the average spread rate at each time since wetting for
  the three constituent species in each mixture. Codes for species in mixtures:
  Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_spread}
\end{figure}

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig8_mixture_obs_vs_pred_consume.pdf}
\caption{Observed vs predicted fuel consumption for litter mixtures. Filled
  circles show observed fuel consumption percentages. Open diamonds indicate
  predicted fuel consumption based on the average fuel consumption at each
  particular time since wetting for the three constituent species in each
  mixture. Codes for species in mixtures: Ab = \emph{Abies concolor}, Ca =
  \emph{Calocedrus decurrens}, Pi = \emph{Pinus jeffreyi} and Qu =
  \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_consume}
\end{figure}

Flammability measures also demonstrated nonadditivity in mixtures. Flame
spread rate was consistently higher than predicted (Fig
\ref{fig:mixture_obs_pred_spread}, Wald test of mixed linear model centered
intercept for observed-predicted values greater than zero, $P$ = 0.017) and the
positive nonadditivity increased with time since wetting (slope greater than
zero, $P$ = 0.005). Fuel consumption, however, showed no overall bias towards
positive or negative nonadditivity, but nonadditivity increased with time
since wetting (Fig. \ref{fig:mixture_obs_pred_consume}, slope not equal to
zero, $P < 0.0001$).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}

\subsection*{Genus-specific moisture dynamics and flammability}

The genera in this study displayed different patterns of both moisture
absorption and desorption, associated with different litter traits. In this
sample of species, those that have a greater absorptive capacity are also among
the most flammable in terms of spread rate when dry. This highlights the
importance of fuel moisture in modulating flammability. One shortcoming of past
work in this system \citep{Magalhaes+Schwilk-2012} was that flammability trials
tested only dry fuels. We therefore expected that rankings of species
flammability might change under more [TODO FIX] realistic moisture levels. For
example, one presumption of ours prior to this experiment was that the high
flammability previously reported for \emph{Q. kelloggii}
\citep{Magalhaes+Schwilk-2012} would disappear under higher fuel moisture
contents. This prediction was partly supported: \emph{Q. kelloggii} fails to
ignite until after 72 hours of drying in this study (Figs.
\ref{fig:spread_time} and \ref{fig:consume_time}). However, high moisture
content in this oak was unable to suppress ignition of mixtures to which it
contributed (Figs. \ref{fig:mixture_obs_pred_spread} and
\ref{fig:mixture_obs_pred_consume}).

The dry down curves we measured demonstrate that after a rain event, the litter
of different species will exhibit different moisture contents over time as a
consequence of both differential initial absorption and of differences in
drying rate. This has consequences for flammability as a function of species
and time since wetting. Our flammability trials demonstrate that both flame
spread rate and fuel consumption increased with time since wetting but that the
rate of increase varied across these taxa. Only the pines are able to ignite
and support flame spread immediately after wetting (Fig.
\ref{fig:spread_time}). In contrast, \emph{Quercus} exhibited a threshold
response to drying with no ignition prior to 72 hours of drying but once
ignition is possible, having faster spread rates than the other taxa. The
Cupressaceae, which in our two species have dense litter but low maximum water
absorption had low flammability across moisture levels. This result is
consistent with the low flammability found in an Australian Cupressaceae that
have small litter particles and dense packing citep{Scarff+Westoby-2006,
  Bowman+Haverkamp-2018}. However, the densely packed litter is not a
consistent trait across the family because \citet{Cornwell+Elvira+etal-2015}
found relatively high litter flammability and low litter density for some Asian
Cupressaceae.

[TODO] [more?]

With eight species, power to detect species-specific trait effects on these
moisture patterns was weak, but we examined one litter trait known to directly
influence flammability: litter bulk density and one leaf trait, specific leaf
area. Litter density is largely driven by leaf size in this system and, in
turn, controls the major axis of flammability variation in dry litter
\citep{Magalhaes+Schwilk-2012}. Our results suggest that moisture dynamics are
controlled by both traits intrinsic to the leaf particles (specific leaf area)
and an emergent property of the litter bed (bulk density). Specific leaf area
here is probably capturing some joint effect of cuticle thickness and total
surface area of litter particles, but we cannot test the morphological and
physical mechanisms of water absorption with these data. The oak has much
higher initial water retention than do the other species, followed by the
finest-needled pine, \emph{P. lambertiana}. Our study examines fresh litter,
but we expect that decomposition will affect moisture dynamics. For example,
Van Wagner \citeyear{Van_Wagner-1969}, investigating fuel weathering, found
that pine needles had a variation in moisture content 3 to 6 times greater when
they lacked their protective wax coating when compared to intact needles.

\subsection*{Effects of moisture and flammability in litter mixtures}

We measured drying rates and flammability of four unique three-species litter
mixtures. These mixtures contained species representative of all dry-down
behavior types observed in single species trials. Mixtures that contained
\emph{Quercus} absorbed more water during wetting, and the one mixture without
any Cupressaceae had the most rapid drying rate (SI Figure 4 and SI Table 8).
Drydown rates, however, did not behave as the average of the constituent
species: mixtures tended to be drier than expected early in the process and
wetter than expected later (Fig. \ref{fig:mixture_obs_pred_mc}). The fact that
mixtures tended to have lower water content than predicted from the mean of the
constituent species after wetting, but higher water content than expected after
long drying is surprising because past work has shown that mixtures tend to
pack less densely than do litter monocultures \citep[][Table
2]{Magalhaes+Schwilk-2012}. If packing density controls rate of drying, then we
would expect mixtures to dry more rapidly than predicted from monoculture
trials. This pattern in moisture content through time, however, did not
directly predict flammability through time.

The overall ranking of flammability in litter mixtures showed some consistency
with the results obtained for the individual species, but there were
significant nonadditive effects. Mixtures containing \emph{Q. kelloggii} and
\emph{P. jeffreyi} tended to have fast spread rates. These results are likely
influenced by the lower bulk densities of litter mixtures containing those two
species. Most interestingly, there were patterns of nonadditivity in spread
rate and in fuel consumption for the four mixtures, indicating that positive
nonadditivity in flammability is not restricted to the completely dry litter
mixtures studied in previous work \citep{Magalhaes+Schwilk-2012,
  VanAltena+Logtestjin+etal-2012}. Nonadditivity in both spread rate and fuel
consumption became more positive with time since wetting and these patterns
were consistent across mixture identities. This is a slightly different result
than that reported by Blauw et al \citeyear{Blauw+Wensink+etal-2015}. Blauw et
al found that nonadditivity was greater in moist mixtures of heathland fuels
than in dry fuels. In their work, nonadditivity was generally negative but
cases of positive nonadditivity in mixtures was often a result of mixtures
meeting ignition thresholds not met by the constituent species alone [TODO:
check wording]. To limit the number of models fit, we limited our investigation
to two flammability parameters that represent two major axes of flammability
variation. Therefore, our study does not fully capture ignition effects
although we do have some results consistent with theirs: mixtures with \emph{Q.
  kellogii} could ignite at times since wetting at which the species in
monoculture could not.

Overall, we report a consistent pattern of increasingly positive nonadditivity
in flammability with litter drying across two largely orthogonal measures of
flammability \citep{Schwilk-2015, Pausas+Keeley+etal-2017,
  Prior+Murphy+etal02018}. This result, together with previously published
work, suggests that litter flammability is prone to nonadditive effects,
suggesting caution in lumping species into fuel categories. Community
flammability is dependent on and particular to its constituent species and
cannot be addressed by simply averaging the flammability of all species
included. In litter-driven fire, positive nonadditivity appears to be the rule,
and the relationships we described here can form a working hypothesis and
baseline expectation to adjust fire behavior predictions in real communities:
in mixtures with fuel moisture content below 50\%, observed flame spread rates
approached double the predicted value based on monoculture trials (Fig.
\ref{fig:mixture_obs_pred_spread}). In completely dry litter from this same
system, past work demonstrates observed spread rates approached three times the
expectation \citep{Magalhaes+Schwilk-2012}.

[TODO rewrite below]

Positive nonadditivity, whereby more flammable species have effects on fire
behavior disproportionate to their contribution to the litter layer, could lead
to vegetation feedbacks. In this temperate mixed conifer forest, these effects
are likely to include positive feedbacks whereby the shade intolerant species
such as the long-needled pines and the large-leaved oak, are favored by the
increased fire to which their fuels contribute \citep{Schwilk+Caprio-2011} and
similar results are reported for southeastern US forests where species
composition shifts have led to decreases in fire activity
\citep{Nowacki+Abrams-2008}. Such positive feedback effects increases the
potential for species to exert community scale effects even when relatively
rare and that can influence community assembly and the evolutionary of
niche-constructing traits \citep{Kerr+Schwilk+etal-1999, Schwilk+Kerr-2002}.
Ongoing climate change and concurrent changes to fire regimes are linked
phenomena. Vegetation-fire feedbacks are important at a wide range of scales
from stand-level species composition shifts to the earth system
\citep{Harris+Remenyi+etal-2016, Archibald+Lehmann+etal-2018}. The patterns we
demonstrate here show the potential importance of species traits and
interactions at local and stand scales where management and fire operations
meet species' biology.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgments}

The authors thank the National Park Service and the employees of Sequoia and
Kings Canyon National Parks. This work was carried out with permission from NPS
under permit \# SEKI-2014-SCI-0040. The authors thank Russell Lackey who
contributed essential assistance with the field work and Gabrielle Plata and
Leticia Plata who assisted with the burning trials. We thank the Lubbock Fire
Department for their ongoing help and cooperation.

This research received no external funding.

\section*{Author contributions}

conceptualization, RMQM and DWS; methodology, RMQM and DWS; experiments, RMQM;
coding and data analysis, DWS and RMQM; writing DWS and RMQM; writing--review
and editing, DWS

\section*{Data and code accessibility}
All data and code for this project is at
\url{https://github.com/schwilklab/trait-flam}. DOI to be provided. For review,
use ``joe'' branch which contains this manuscript and code needed to create the
figures and tables.

\newpage
\bibliography{Magalhaes_Schwilk_litter_flam_moisture.bib}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \conflictsofinterest{The authors declare no conflict of interest. The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results.}

% References, variant B: external bibliography
% =====================================

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}