### 2019-12-20 ms/Magalhaes_Schwilk_litter_flam_moisture.tex

%% Revision 2019-12-10
\documentclass[letterpaper,12pt]{article}
\usepackage{jecol}

\title{Moisture absorption and drying alters nonadditive litter flammability
  in mixed conifer forest}

\running{Litter flammability}

\author{Rita Quinones de Magalhaes$^{1,2}$ and Dylan W. Schwilk$^1$}

\affiliations{
    \item Texas Tech University, Lubbock, Texas 79411, USA.
    \item Current address: Rochester Institute of Technology, Rochester, NY 14623, USA.
}

\nwords{8200}
\ntables{4}
\nfig{8}
\nref{55}

\corr{\url{dylan.schwilk@ttu.edu}}

\begin{document}
\maketitle

\begin{abstract}
  \noindent \begin{enumerate}

      \item Moisture content is a strong determinant of forest fuel
    flammability, however, little is known about the mechanisms by which
    species-specific litter traits influence litter moisture absorption and
    retention. Characterizing how litter fuels retain moisture through time and
    how fuels from multiple species interact are key steps to adding species or
    taxon specific information to fire behavior prediction.

      \item To fill these gaps in knowledge, we determined moisture absorption
    and retention characteristics of the litter from multiple tree species,
    investigated what litter traits influenced these processes, and tested for
    nonadditive moisture dynamics and nonadditive flammability in multi-species
    mixtures. The litter from eight tree species of a mixed-conifer forest was
    saturated and allowed to dry to determine moisture absorption capacity and
    drying rates. Burn trials were performed in litter beds of single species
    and of mixtures to establish flammability response to the dry-down process.

      \item Moisture dynamics varied across species with higher specific leaf
    area being associated with higher maximum absorption and lower litter bulk
    density being associated with faster drying rates. We examined two axes of
    flammability variation and found that these differing moisture dynamics
    result in time since wetting having strongly differing effects on
    flammability across the major litter types typified by different tree taxa.

    \item Litter mixtures exhibited nonadditivity in flame spread rate and
    in fuel consumption. This nonadditivity became more positive with fuel
    drying.

      \item \emph{Synthesis:} Taxon specific patterns of moisture absorption
    and drying can change flammability rankings under moist compared with dry
    fuels because traits that increase flame spread rate increase water
    absorption. Litter mixtures generally burn with fire behavior more similar
    to that of the most flammable constituent species and this effect increases
    as fuels dry.

  \end{enumerate}
\end{abstract}

\noindent \textbf{Keywords:} California, fire, flame spread rate, flammability,
fuel consumption, fuel moisture, leaf litter, nonadditivity

\newpage

\section*{Introduction}

Fuel moisture content is an important, and often driving, determinant of
surface fire behavior \citep{Rothermel-1972}. Moisture acts as a heat-sink
resulting in less energy available for propagation, and the resulting water
vapor will both cool the flaming front and dilute flammable gases
\citep{Albini-1976, Shafizadeh-1977}. Environmental conditions such as
temperature, relative humidity, and wind drive the moisture content of the fuel
available to burn \citep{Kreye+Varner+etal-2018}. Variation in moisture of
living fuels has strong effects on flammability, but moisture of dead excised
litter fuels can also vary. In particular, fine dead fuels respond very readily
to environmental changes \citep{Nelson-2001}. Moisture of litter fuels varies
with climate, with seasons, and with short term weather within a season.
Therefore, both the rate of drying and flammability as a function of moisture
content are key parameters that influence fuel flammability throughout the fire
season.

Flammability is a multi-dimensional characteristic of fuels
\citep{Schwilk-2015, Pausas+Keeley+etal-2017} with different traits influencing
different components of fire behavior. In many forests that experience surface
fire, litter properties influence fire behavior
\citep{Ganteaume+Marielle+etal-2011, Schwilk+Caprio-2011,
  Varner+Kane+etal-2015}. In litter driven fire, the two major axes of
variation are described by measurements related to flame spread rate and
measurements related to heat release or heat transfer
\citep{Magalhaes+Schwilk-2012, Prior+Murphy+etal-2018}. These different
components of fire behavior have differing ecological effects with spread rate
being associated with rates of heat release and area burned, and with heat
transfer predicting soil impacts of fire and tree mortality. A shortcoming in
our current knowledge is our integration of specific species and plant traits
into fuel hazard assessment, fire modeling, and fire effects prediction.
Developing quantitative yet robust and general ways to integrate actual plant
communities into fuel assessment will require understanding the magnitude of
flammability variability across species in space and time and parameterizing
based on plant characteristics.

Species differ in their litter flammability \citep{Fonda+Belanger+etal-1998,
  Fonda-2001, Scarff+Westoby-2006, Kane+Varner+etal-2008,
  Cornwell+Elvira+etal-2015} and fire regime can vary over fine spatial scales
in response to community composition and species-specific flammability
\citep{Stephens-2001, Hille+Stephens-2005, Schwilk+Caprio-2011}. Litter
packing, measured as litter-bed density or packing ratio, is an important
control on ignition and spread in models of surface fire \citep{Weber-1991,
  Balbi+Santoni+etal-1999, Morvan+Larini-2001}. Multiple studies have
demonstrated that litter bed density controls the flammability of dry litter
fuels in seasonally dry climates where higher litter density reduces flame
spread rate. Across these studies, there is increasing consensus that larger,
less densely packed litter beds increase oxygen availability and burn with
faster spread rates \citep{Scarff+Westoby-2006, Schwilk+Caprio-2011,
  Magalhaes+Schwilk-2012, Cornwell+Elvira+etal-2015, Varner+Kane+etal-2015,
  Kreye+Varner+etal-2018}. The litter packing is driven by litter particle size
and shape, which is in turn influenced by leaf size, how leaves or branches are
abscised, and subsequent decomposition \citep{Fonda+Belanger+etal-1998,
  Stephens+Finney+etal-2004, Scarff+Westoby-2006, Kane+Varner+etal-2008,
  Cornwell+Elvira+etal-2015, Zhao+Cornwell+etal-2016,
  Grootemaat+Wright+etal-2017, Prior+Murphy+etal-2018,
  Zhao+vanLogtestijn+etal-2019}.

These species-specific flammability effects are increasingly well established.
However, how moisture dynamics modulate species specific flammability is poorly
known. Fuel moisture often limits fire activity in the western US
\citep{Miller+Urban-1999, Parks+Parisien-2014}. Fire scientists categorize fuel
by size classes that correspond to moisture exchange rates, although such
simplified categories have long been recognized as not sufficient to describe
species-specific effects on moisture dynamics \citep{Anderson-1985} which are
dependent upon forest structure (shading and rain through-fall effects), litter
physical structure, and chemical composition of the fuel-beds
\citep{Nelson+Hiers-2008, Matthews-2014, Kreye_Hiers_etal-2018}. A variety of
models describing moisture dynamics of fine fuels are used by fire scientists
\citep{Viney-1991, Nelson-2000, Catchpole+Catchpole+etal-2001}. These models
usually aim to describe equilibrium conditions and at least include temperature
and relative humidity, although some include precipitation and solar radiation
effects as well. Explicit models of fire behavior as a function of time since
wetting could improve fire prediction, however. Long term drying following snow
melt will influence the length of the fire season. Short term drying following
rainfall during the fire season will influence the probability of sustained
ignition, spread rates, and fuel consumption when ignition occurs. That species
can differ in litter drying rates has been demonstrated for at least one forest
system \citep{Kreye+Varner+etal-2013}.

Differential drying is especially important when precipitation occurs
during the same season as ignitions. Although synoptic weather conditions that
create dry and wet lightning storms differ, the potential for precipitation and
lightning during the same overall period of thunderstorm activity often exists:
wet and dry convection days can occur under overlapping conditions,
\citep{Rorig+Ferguson-1999}. Moisture levels of fine fuels respond rapidly to
precipitation events during the fire season \citep{Estes+Knapp+etal-2012}.
Differential moisture content across time may contribute to the often high
spatial heterogeneity of surface fires \citep{Hille+Stephens-2005,
  Knapp+Keeley-2006, Kreye_Hiers_etal-2018} and to variation in fire regime
across landscapes \citep{Stephens-2001}.

Fuel moisture content through time is controlled by two processes, wetting
(including absorption and adsorption) and drying rate (desorbtion)
\citep{Anderson-1990, Kreye+Kobziar+etal-2013}. Absorptive capacity represents
the potential maximum fuel moisture and is evaluated by the moisture content of
litter when saturated; hereafter we refer to this measurement as ``maximum
water absorption''. Differences in moisture content over time since saturation
can indicate different abilities to retain moisture. Leaf morphology,
especially leaf cuticle thickness, can influence water absorption
\citep{Van_Wagner-1969}. Litter particle geometry may have effects on maximum
moisture content because a larger total surface area will increase water
adsorption. Drying rate should also be influenced by leaf and litter bed traits
because surface-area-to-volume ratio and litter bulk density will influence air
flow through the litter bed. Therefore, we expect litter packing to effect fire
behavior in two ways: directly by its effect on oxygen availability for
combustion, and indirectly by its effect on drying rate. Both these pathways
are affected by leaf particle traits, especially leaf size
\citep{Scarff+Westoby-2006}. Fuel-beds composed of larger leaves or leaves with
thin cuticles should favor moisture absorption. On the other hand, large leaves
lower litter density and should favor drying, thus increasing the probability
of ignition and fire spread. The interaction of mositure absorption and drying
deserves investigation.

In mixed species communities, species traits may interact such that the
flammability of multi-species mixtures is not the average of the constituent
species. Such nonadditive effects are often synergistic in dry litter fuels
where the flammability of a mixture is driven by the most flammable species
\citep{VanAltena+Logtestjin+etal-2012, Magalhaes+Schwilk-2012}, but negative
nonadditivity has also been recorded for litter mixtures
\citep{Blauw+Wensink+etal-2015, Zhao+vanLogtestijn+etal-2019}. It is possible
that similar nonadditive effects may influence litter moisture content and
flammability of moist fuels. There is some evidence for nonadditive
flammability in moist fuels with one study finding increasing nonadditivity
with moisture \citep{Blauw+Wensink+etal-2015}. In mixed conifer forest in
California, where the fire regime is characterized by surface fires fueled by
plant litter, overstory tree composition drives variation in local fire
behavior \citep{Schwilk+Caprio-2011} as a result of differential litter
flammability across tree species and as a result of synergistic interactions in
multi-species litter mixtures \citep{Magalhaes+Schwilk-2012}. These
flammability feedbacks may help explain historical fire frequency patterns
\citep{Schwilk+Caprio-2011}, but current estimates of litter flammability in
mixed conifer forests are based largely on burning dry litter.

We asked if differences in litter moisture absorption and drying rates could
change the relative flammability of litter across species compared with
flammability measurements based on dry fuels. We investigated these litter
moisture dynamics across eight tree species in a temperate fire-prone forest in
California. We determined the maximum water absorption and rate of moisture
loss and examined how these traits vary in relation to specific leaf area and
litter packing density. Experimental burn trials of leaf litter at various
moisture levels determined how time since wetting influenced two axes of
flammability characterized by flame spread rate and total fuel consumed. To
determine the potential magnitude of mixture effects, we tested if the
nonadditive effects on reported in previous studies occur under higher moisture
levels, and whether fuel moisture effects strengthen or moderate any
nonadditivity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Materials and Methods}

\subsection*{Site description and species selection}

Field sites for this study were located in Sequoia and Kings Canyon National
Parks, California, USA (N$36^\circ 36'$, W$118^\circ 42'$) between 1600 and
2400\,m elevation in mixed-conifer forest. We chose eight dominant tree species
that are representative of this type of forest: \emph{Pinus jeffreyi} Grev. \&
Balf., \emph{Pinus lambertiana} Dougl., \emph{Pinus ponderosa} Dougl. ex Laws.,
\emph{Abies concolor} (Gord. \& Glend.) Lindl. ex Hildebr., \emph{Abies
  magnifica} A. Murr., \emph{Calocedrus decurrens} (Torr.) Florin,
\emph{Quercus kelloggii} Newb., and \emph{Sequoiadendron giganteum} (Lindl.) J.
Buchholz. In this study, we consider litter the top layer of leaves and small
twigs less than 1\,cm diameter that have fallen in the previous year (mostly
undecomposed or lightly decomposed litter representing fuels classified as
1-hour and 10-hour fuels). Leaf litter was collected in the summer (mid-June to
mid-July) in 2012 from 21 separate sites across the study area. We collected
from a minimum of four populations per species with each population within a
species was 9--32\,km distant from its neighbors. An exception to this was
\emph{A. magnifica} which is restricted to the higher elevations and could only
be collected from one area due to designated wilderness boundaries and
logistical constraints. Within that area, \emph{A. magnifica} litter was
collected from over 10 different individuals. For the remaining species, the
collection involved 2--4 individual trees at each site which were separated
from one another by at least 10\,m. The litter was collected 2\,m away from the
trunk to obtain a more uniform sample of leaf and fine branchlet litter,
because bark and heavier branches tend to fall closer to the tree trunk. The
top 2.5-5\,cm of litter were collected by raking and shoveling into large bags.
All litter was oven dried to $< 5$\% fuel moisture by dry weight, then weighed
and divided into moisture trials. The collected litter was stored in large
paper bags and cardboard boxes in a lab storage space. We minimized handling
and moving the bagged litter to avoid mechanical damage and changes to the
particle size distributions. The moisture manipulations and subsequent burning
trials were conducted in April--August 2015.

\subsection*{Moisture absorption and desorption}

We produced dry-down curves to determine maximum water absorption and the rate
of drying for each species and for four distinct three-species mixtures. These
curves allowed us to determine the rate at which a species loses moisture,
approximated as exponential decay, and to determine maximum water content
following saturation and gravity draining. Samples from the eight species (six
replicates per species) and four mixtures (5 replicates per mixture type) were
placed in baskets approximately 45 x 45\,cm by 15\,cm tall constructed of
aluminum mesh with 0.52\,cm square openings. Samples were 450\,g dry weight
litter per trial and litter depth was standardized to 10\,cm by changing the
horizontal dimensions of the basket. Weight was measured by using a balance
sensitive to 0.1\,g (model XS16001L, Mettler Toledo, Columbus, OH). Samples
were immersed in plastic storage bins filled with water for 24 hours to fully
saturate them. After saturation, the baskets were removed from the water,
allowed to drain for 3 minutes, and samples weighed to measure maximum moisture
content. The samples were then allowed to dry in a controlled environment kept
at $21^\circ$C temperature and 30\% relative humidity, assessed via an iButton
sensor (Maxim Integrated) placed 1\,m above the samples. Litter baskets were
arranged in a completely randomized design. During the dry-down process three
subsamples of 5--10\,g were taken at 24 hour intervals and weighed. These
subsamples were then oven dried for 24 hours at $100^\circ$C and re-weighed to
assess moisture content. This was repeated until the final sample from a basket
was less than 10\% moisture. This experimental setup was designed to accelerate
drying compared with field conditions because the baskets were suspended in air
2.8 -- 5.3$^\circ$C above the warmest month average temperatures for the
closest weather stations to the collection sites (Grant Grove and Lodegpole
stations, NOAA 1981-2010 normals,
https://www.ncdc.noaa.gov/cdo-web/datatools/normals).

With only eight species in this study, our power to test leaf trait effects on
the observed moisture dynamics was low, so we decided \emph{a priori} to test
the effect of one leaf trait likely to influence moisture absorption and
adsorption (specific leaf area), and one litter bed trait known to directly
affect flammability and likely to influence drying rate (litter bulk density).
Specific leaf area was measured on freshly fallen litter by sampling 5
individuals per species. This measurement included the petiole and for the
scale-leaved species included the full excised branchlet. Litter bulk density
was obtained in this experiment for each basket by dividing the dry mass
(450\,g) by the estimated litter volume (basket width x length x litter depth).
To examine litter particle size distributions in the stored litter, we
subsampled 10 approximately 2\,g samples per species from the stored litter.
Each individual particle of that subsample was measured for length (longest
axis), width (longest axis perpendicular to length) and thickness
(perpendicular to length and width at midrib) using an electronic caliper.

\subsection*{Flammability assessment}

We performed flammability tests across multiple moisture levels on monospecific
litter beds and on mixed litter beds with each species or mixture replicated 5
times at each moisture level. We used the initial dry-down experiment to guide
our timing so as to attempt to burn samples over a range of moisture contents.
We staggered saturation and start of drying so we could conduct our burning in
a blocked design with multiple species at the same moisture content per burning
day. We aimed for moisture contents from 10\% to 100\% on a dry mass basis because
preliminary trials on damp fuels suggested that across these fuels ignition
will not occur above 100\% fuel moisture. To expand our data on the dry end of
this gradient, we supplemented these data with data from an earlier experiment
on oven-dried litter \citep{Magalhaes+Schwilk-2012}. This supplemental data
added five replicates per species at an oven dried moisture content of $<$ 5\%.
We assigned these replicates a time since wetting value of 144 hours equivalent
to the longest time since drying for the wetted samples.

Burn trials occurred in a cement structure used to simulate house fires at the
Fire Department of the City of Lubbock. This structure minimized wind and
helped regulate temperature and relative humidity. Temperature and relative
humidity were measured before every trial using a Kestrel 3000
(Nielsen-Kellerman, Boothwyn, PA) and these values were used to calculate vapor
pressure deficit as a single weather covariate at each trial. All trials were
conducted on clear days between 10 am and 3 pm. We scheduled burns in a
randomized blocked design aiming to burn one replicate of each species or
mixture on single burn day. This required scheduling species saturation and
during start based on the dry down curves.

Flammability was assessed using a 150\,cm long burn table, 15\,cm wide and
15\,cm tall, in which leaf litter was placed and the table gently shaken to
allow settling. Burning trials all used 450\,g (dry weight) of litter. For
three-species mixtures, this was 150\,g per species. Therefore, fuel mass and
burning tray length and width were fixed, but fuel depth was allowed to vary.
We recorded the actual fuel depth at three points along the tray to calculate
litter bed density before burning. This design allows fire to reach constant
flame spread rates \citep{Magalhaes+Schwilk-2012}. Two metal rulers equally
spaced along the apparatus allowed visual determination of maximum flame
height. Litter was ignited with a propane torch; maximum adiabatic flame
temperature of 1899$^\circ$C,
\url{http://www.benzomatic.com/products/fuel.aspx}) at one end of the apparatus
and allowed to burn until extinction. If a sample failed to ignite after 30
seconds, no further attempt was made and if three samples of a species at a
particular moisture level failed to ignite, no further attempts were made on
that species/moisture combination and the replicate was allowed to continue
drying until the next scheduled checkpoint. The flammability trials followed
the methodology detailed in \citet{Magalhaes+Schwilk-2012}.

During flammability trials, we recorded time to ignition (s), duration of
flaming combustion (s), flame spread rate (mm/s), calculated by dividing the
length of the burned surface by the time it takes the flaming front to reach
the end of the apparatus, maximal flame height (mm), and fuel consumption (\%).
Some trials resulted only in smoldering combustion; for those trials, a time to
ignition and fuel consumption were recorded, but flame spread rate and flame
height were recorded as zero. Our thermocouple system failed or data was lost
for two days of burning trials, therefore we have omitted those temperature
data from this analysis. Our past work \citep{Magalhaes+Schwilk-2012}
demonstrated that flame spread rate captures one important axis of flammability
variation \citep{Schwilk-2015, Pausas+Keeley+etal-2017, Prior+Murphy+etal-2018}
and percent fuel consumed captures a portion of the second major axis of
flammability variation. Our analysis, therefore, focused on these two
relatively orthogonal measures of flammability to avoid examining multiple
redundant dependent variables.

\subsection*{Statistical analyses}

Dry-down curves were created based on the drying experiment, fitted with an
exponential decay curve to each species:

\begin{equation}
m = m_{max} e^{-\lambda t}
\end{equation}

Where $m$ is moisture content at time $t$, and $\lambda$ is the exponential
decay coefficient. To estimate these coefficients we fit a linear mixed-effects
model by first taking the natural logarithm of moisture content. Taxonomic
group was a fixed effect and time since wetting was a fixed effect nested
within replicate basket. We extracted the coefficients of the fitted model to
describe maximum water absorption ($m_{max}$; g of water per g of dry mass) and
drying rate ($\lambda$; hour$^{-1}$) for each species. All analyses were
conducted in R 3.6.1 \citep{RCoreTeam-2019}. We fit models with R packages
`lme4' \citep{Bates_Machler_etal-2015} and `afex'
\citep{Singmann_Bolker_etal-2017}. Approximate degrees of freedom and p-values
were calculated by the Kenward-Roger approximation \citep{Kenward_Roger-1997}
which is recommended by Luke \citeyear{Luke-2017} as producing acceptable Type
I error rates. We tested if one litter particle trait (specific leaf area) and
one litter bed trait (litter bulk density) influenced maximum water absorption
and drying rate by fitting linear models with species means as observations.
For each response variable we fit a single linear model with two predictors and
an interaction term.

Our past work \citep{Magalhaes+Schwilk-2012} demonstrated that dry litter
flammability across these eight species was driven by leaf traits and litter
packing. These previously reported flammability trials demonstrated that
species within a genus tended to behave similarly and the two scale-leaved
genera in the Cupressaceae, \emph{Calocedrus} and \emph{Sequoiadendron}, also
behaved similarly to one another. Therefore, to preserve power we used
taxonomic groups (\emph{Abies}, Cupressaceae, \emph{Quercus}, and \emph{Pinus})
as our flammability functional groups and treated these as fixed effects in our
models. The effect of fuel moisture and time since wetting on flammability
parameters were modeled using linear models with fuel moisture or time since
wetting and taxon as fixed effects and average vapor pressure deficit in the
burning room during that trial as a nuisance covariate. We used linear mixed
models to predict spread rate, but fuel consumption tends to exhibit threshold
effects with few intermediate values. Therefore, we used mixed generalized
linear models with a beta regression link with the `glmmADMB' package
\citep{Skaug_Fournier_etal-2016} to model percent fuel consumed as a function
of moisture content and of time since wetting.

We investigated possible nonadditive effects on moisture and on flammability by
testing if moisture content and flammability of mixtures were predicted by the
average of the constituent species. To choose mixture combination, we first
grouped the species according to their maximum water absorption and drying rate
according to estimated marginal means using the R package `emmeans'
\citep{Lenth-2019}. This resulted in four groups of species and we then
selected one representative species from each litter type group and produced
all four possible three-species litter mixtures using those four species. For
each mixture and time since wetting, we predicted moisture content based on the
mean moisture content of the three constituent species at that time since
wetting. We predicted flammability parameters based on the average flammability
measures from individual species trials at each time since wetting and then
calculated a predicted spread rate and fuel consumption as the average of these
three values for the mixture. The expectation from the null model was that the
difference between observed and predicted values of flammability will be zero.
We fit linear mixed effect models with the observed-predicted flammability
value as the dependent variable and used Wald tests to test if the centered
intercept was different than zero (a measure of overall nonadditivity) and if
the slope with time since wetting was different from zero (indicating
nonadditivity changing with drying).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}

\subsection*{Moisture dynamics across species}

Past work has demonstrated that green leaf length predicts litter packing and,
consequently, flame spread rate across these species. Litter particle size
distributions, however show that there is substantial variation in particle
sizes within a species (SI Fig.~1). There were differences in the maximum water
absorption (all $P < 0.0001$, Fig.~\ref{fig:drydown} and
Table~\ref{tab:drydown}). \emph{Q. kelloggii}, the only broadleaf, is the
species that absorbed the most moisture initially. \emph{Quercus}, like the
pines, had a rapid drying rate (Fig.~\ref{fig:drydown} and SI Fig.~2).
Calculating estimated marginal means following the linear mixed-effects model
discriminated the species into three different maximum water absorption groups:
1) \emph{Q. kelloggii}; 2) \emph{P. lambertiana} and \emph{C. decurrens}; and
3) \emph{A. magnifica}, \emph{S. giganteum}, \emph{A. concolor}, \emph{P.
  jeffreyi} and \emph{P. ponderosa} (SI Fig.~2). There are also differences in
drying rates across the eight species and estimated mixed linear model slopes
discriminated species into three groups from lowest to most rapid: 1) \emph{C.
  decurrens}, 2) \emph{Abies spp} and \emph{S. giganteum}, and 3) \emph{Q.
  kelloggii} and \emph{Pinus spp} (SI Fig.~3).

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig1_drydown-curves.pdf}
\caption{Dry down curves for eight litter types. Lines show best-fit
  exponential decay curves fit to moisture content on a dry mass basis. Litter
  samples we saturated by immersion for 24 hours and then allowed to drain by
  gravity for 3 min before initial weighing. Colors indicate genera and there
  were significant differences in both maximum water absorption and in rate of
  drying (based linear model on log of water content). A small amount of
  horizontal jitter (0.5 hr) was added to this plot to aid in visualizing
  because, although collections were staggered, there were sometimes multiple
  samples collected at a single time point.}
 \label{fig:drydown}
\end{figure}

\begin{table}[h]
  \caption{Linear mixed model results for moisture content as a function of
    time since wetting and species. Approximate degrees of freedom (df), pseudo F
    statistics and p-values were calculated by the Kenward-Roger approximation.
    \citep{Kenward_Roger-1997}. Model coefficients reported in SI Table~1.}
  \label{tab:drydown}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{llll}
  \toprule
Effect & df & $F$ & $P$ \\ 
  \midrule
  hour & 1, 183.73 & 1599.59 & $<$.0001 \\ 
  species & 7, 89.73 & 36.15 & $<$.0001 \\ 
  hour:species & 7, 184.15 & 23.06 & $<$.0001 \\ 
   \bottomrule
\end{tabular}
\end{table}

Across these eight species specific leaf area had a positive association with
maximum water absorption (Table~\ref{tab:mc_di_anova}, SI Fig~5, linear model
$P = 0.003$) and litter bulk density had a negative effect on drying rate
(Table~\ref{tab:mc_di_anova}, SI Fig~6, linear model $P = 0.029$).

\begin{table}[h]
  \caption{Linear model results for maximum water absorption (\% moisture
    content on dry mass basis) and for estimated drying rate (hr$^{-1}$) as
    functions of specific leaf area and litter bulk density across eight
    species. ANOVA tables shown with degrees of freedom, sum of squares, mean
    squares, $F$, and $P$ for each effect. Model coefficients in SI Tables~2 and 3}
  \label{tab:mc_di_anova}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.

\begin{tabular}{ccrrrrr}
  \toprule
Ind. variable & Effect & df & Sum of Squares & Mean squares & $F$ & $P$ \\ 
  \midrule
Moisture content & SLA & 1 & 250834.5 & 250834.51 & 130.13 & 0.000 \\ 
  & density & 1 & 538.5 & 538.54 & 0.28 & 0.625 \\ 
  & SLA:density & 1 & 5135.2 & 5135.24 & 2.66 & 0.178 \\ 
  & Residuals & 4 & 7710.4 & 1927.61 &  &  \\ 
  \midrule

Drying rate & density & 1 & $3.92 \times 10^{-4}$ & $3.92 \times 10^{-4}$ & 11.01 & 0.029 \\ 
 &  SLA & 1 & $1.88 \times 10^{-5}$ & $1.88 \times 10^{-5}$ & 0.53 & 0.507 \\ 
 &  density:SLA & 1 & $6.89 \times 10^{-6}$ & $6.89 \times 10^{-6}$ & 0.19 & 0.683 \\ 
 &  Residuals & 4 & $1.42 \times 10^{-4}$ & $3.56 \times 10^{-5}$ &  &  \\ 
   \bottomrule

\end{tabular}
\end{table}

\subsection{Flammability across taxonomic groups}

Flame spread rate declined with fuel moisture (Fig.~\ref{fig:spread_moist}).
\emph{Quercus} had higher spread rates across all fuel moistures and had a
steeper decline in response to moisture than did the other groups (taxon by
time since wetting interaction $P = 0.022$, Table~\ref{tab:spreadrate_anova}).
Fuel moisture negatively influenced fuel consumption. Fuel consumption and the
rate of change with moisture content varied across the taxonomic groups
(Fig.~\ref{fig:consume_moist}, Table~\ref{tab:consume_anova}). Although the
overall ranking of groups was similar for fuel consumption and for flame spread
rate, there were different pattern across taxa in these two flammability
parameters. For spread rate, the decline in flammability with moisture was
shallowest for \emph{Pinus} and Cupressaceae (Fig.~\ref{fig:spread_moist}). For
fuel consumption, \emph{Pinus} and \emph{Abies} has a steeper decline in
flammability with fuel moisture than did the other groups
(Fig.~\ref{fig:consume_moist}).

%% spread rate by moisture
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig2_spread_actualMC.pdf}
\caption{Flame spread rate by moisture content (model fit $F = 61.15$, df = 9,
  148, $P < 0.0001$). Model ANOVA table in Table \ref{tab:spreadrate_anova} and
  coefficients in SI Table 4.}
  \label{fig:spread_moist}
\end{figure}

%% Table 3 
\begin{table}[h]
  \caption{Generalized linear mixed model results for two models predicting
    flame spread rate: the first model examined spread rate as a function of
    fuel moisture (mc) with taxonomic group as a fixed effect and species as a
    nested random effect. The second model predicted spread rate as a function
    of time since wetting (time) with taxonomic group as a fixed effect and
    species as a nested random effect. Approximate degrees of freedom (df),
    pseudo F statistics and p-values were calculated by the Kenward-Roger
    approximation \citep{Kenward_Roger-1997}. Estimated coefficients are in SI
    Tables 4 and 6}
  \label{tab:spreadrate_anova}

\centering
\begin{tabular}{ccrrrrr}
  \toprule
Model set & Effect & df & $F$ & & $P$ \\
  \midrule
  Effect of moisture & mc & 1, 146.20 & 21.04 & $<$.0001 \\ 
 & taxon & 3, 51.58 & 65.08 & $<$.0001 \\ 
 & vpd & 1, 148.79 & 9.62 & .002 \\ 
 & mc:taxon & 3, 147.21 & 4.83 & .003 \\

  \midrule

  Effect of time & hour & 1, 214.50 & 66.22 & $<$.0001 \\ 
 &  taxon & 3, 7.59 & 61.81 & $<$.0001 \\ 
 & vpd & 1, 215.00 & 13.49 & .0003 \\ 
 & hour:taxon & 3, 183.97 & 3.99 & .009 \\

   \bottomrule

\end{tabular}
\end{table}

%% fuel consumption by moisture:
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig3_consume_actualMC.pdf}
\caption{Percent fuel consumed by moisture content. Lines indicate best fit
  beta regression lines. Model deviance table in Table \ref{tab:consume_anova}
  and coefficients are in SI Table 5.}
  \label{fig:consume_moist}
\end{figure}

\begin{table}[h]
  \caption{Generalized linear mixed model results for two sets of models
    predicting fuel consumption. The first set modeled fuel consumption as a
    function of fuel moisture (mc) with taxonomic group as a fixed effect and
    species as a nested random effect. The second set modeled fuel consumption
    as a function of as a function of time since wetting (time) with taxonomic
    group as a fixed effect and species as a nested random effect. The analysis
    of deviance table below shows sequential chi-square tests on nested beta
    regression models: from the simplest null model within each group with only
    intercept and random effect, for a taxon effect only model, for a taxon and
    moisture content effect, and for the full model including the taxon by
    moisture or time interaction. Coefficient estimates are in SI Tables 5 and
    7.}
  \label{tab:consume_anova}
  \centering

  \begin{tabular}{llrrrr}
\toprule
Model set & model & Log Lik. & df & Deviance & $P$ \\ 
\midrule
  Effect of moisture &   \~{} 1 & 262.12 &  &  &  \\ 
   &  \~{} taxon & 268.73 & 3 & 13.23 & 0.0042 \\ 
   & \~{} taxon + mc & 289.57 & 1 & 41.68 & $<$ 0.0001 \\ 
   & \~{} taxon + mc + taxon:mc & 293.87 & 3 & 8.60 & 0.0351 \\ 
\midrule
Effect of time &  consum \~{} 1 & 140.08 &  &  &  \\ 
 & \~{} taxon & 155.84 & 4 & 31.51 & $<$ 0.001 \\ 
 & \~{} taxon + time & 209.84 & 1 & 108.01 & $<$ 0.001 \\ 
 & \~{} hour + taxon + taxon:time & 216.96 & 4 & 14.23 & 0.01 \\ 
    \bottomrule
    \end{tabular}
\end{table}

As a consequence of the relationships between fuel moisture and flammability,
time since wetting had a significant effect on spread rate ($P < 0.0001$) and
the slope of this effect varied by taxonomic group (Fig. \ref{fig:spread_time},
Table~\ref{tab:spreadrate_anova}). The pines were able to ignite and burn much
sooner after wetting than did the other taxa, none of which had measurable
spread rates until burning was attempted 72 hours after wetting. Trials that
failed to even ignite smoldering combustion are omitted completely from the
model and figure. At 72 hours, the \emph{Abies}, \emph{Calocedrus}, and
\emph{Sequoiadendron} had dryer litter than did the \emph{Quercus}, yet the
spread rates were slow. \emph{Quercus} had the steepest response to time since
wetting (Fig.~\ref{fig:spread_time}, SI Table~6)).

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig4_spread_time.pdf}
\caption{Flame spread rate by time since wetting for litter of eight species.
  Lines indicate best fit linear model fits. Model ANOVA table in Table
  \ref{tab:spreadrate_anova} and coefficients in SI Table 6. Horizontal jitter
  (up to 2 hr) was added to the plot for visualization and to avoid
  overplotting. All collections were at the times labeled on the x-axis. A lack
  of data at earlier time points for a taxon indicates no ignitions occurred at
  those time points.}
\label{fig:spread_time}
\end{figure}

Fuel consumption was highly variable across trials and intermediate values were
less common than were very low or moderately high consumption
(Fig.~\ref{fig:consume_time}). All parameters significantly contributed to the
generalized linear mixed beta regression model as indicated by the nested model
comparisons in Table~\ref{tab:consume_anova}. As was the pattern for flame
spread rate, pines had rapid flammability increases with drying. However,
unlike in the other flammability analyses, here the pines did not all behave
similarly: \emph{P. lambertiana} had lower fuel consumption than did \emph{P.
  ponderosa} and \emph{P. jeffreyi}. We noticed this during trials and in model
fitting we found that placing \emph{P. lambertiana} in its own group allowed
model convergence that was otherwise impossible. This was at odds with our
intent to use only \emph{a priori} taxonomic groups, but we could not obtain
model convergence otherwise. This result should therefore be seen as
preliminary because the test does not reflect an \emph{a priori} hypothesis,
but probably reflects the fact that \emph{P. lambertiana} has shorter and finer
needles than the other two pines.

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig5_consume_time.pdf}
\caption{Percent fuel consumed by time since wetting for litter of eight
  species. Curves shown indicate best fit beta regression lines. Model ANOVA
  table in Table \ref{tab:consume_anova} and coefficients in SI Table 7.
  Horizontal jitter (up to 2 hr) was added to the plot for visualization and to
  avoid overplotting. All collections were at the times labeled on the x-axis..
  A lack of data at earlier time points for a taxon indicates no ignitions
  occurred at those time points.}
  \label{fig:consume_time}
\end{figure}

\subsection*{Nonadditivity in  litter mixtures}

Based on visual inspection of the dry down curves and the estimated marginal
means for maximum water absorption and drying rate (SI Figs 3--4), we selected
four species to represent the full range of litter moisture dynamics for use in
the litter mixtures and examined all four three-species combinations of those
four. The four species were \emph{Abies concolor}, \emph{Calocedrus decurrens},
\emph{Pinus jeffreyi} and \emph{Quercus kelloggii}.

Mixture drying rates and maximum water absorption showed some relationship to
the dry down curves for individual constituent species (see SI Fig.~7). For
example, the mixture that does not contain \emph{Q. kelloggii} (AbCaPi in SI
Fig.~7) has the lowest maximum water absorption. Maximum water absorption and
drying rates differed across mixtures (linear model based on log of moisture
content, SI Tables~8 and 9).

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig6_mixture_obs_vs_pred_mc.pdf}
\caption{Observed vs predicted fuel moisture content for litter mixtures.
  Predictions are based on the mean moisture content at each time since wetting
  for the three constituent species in each mixture. Codes for species in
  mixtures: Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_mc}
\end{figure}

There was evidence for nonadditive effects in drying rate. Mixtures tended to
have lower water content than predicted from the mean of the constituent
species after wetting, but had higher water content than expected after long
drying (Wald test shows slope of mixed linear model
[Fig.~\ref{fig:mixture_obs_pred_mc}] less than 1, $P < 0.0001$).

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig7_mixture_obs_vs_pred_spread.pdf}
\caption{Observed vs. predicted flame spread rate for litter mixtures. Filled
  circles show observed spread rate values. Open diamonds indicate predicted
  spread rates based on the average spread rate at each time since wetting for
  the three constituent species in each mixture. Codes for species in mixtures:
  Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_spread}
\end{figure}

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig8_mixture_obs_vs_pred_consume.pdf}
\caption{Observed vs predicted fuel consumption for litter mixtures. Filled
  circles show observed fuel consumption percentages. Open diamonds indicate
  predicted fuel consumption based on the average fuel consumption at each
  particular time since wetting for the three constituent species in each
  mixture. Codes for species in mixtures: Ab = \emph{Abies concolor}, Ca =
  \emph{Calocedrus decurrens}, Pi = \emph{Pinus jeffreyi} and Qu =
  \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_consume}
\end{figure}

Flammability measures also demonstrated nonadditivity in mixtures. Flame spread
rates were generally higher than predicted (Fig.
\ref{fig:mixture_obs_pred_spread}, Wald test of mixed linear model centered
intercept for observed-predicted values greater than zero, $P = 0.017$) and the
positive nonadditivity increased with time since wetting (slope greater than
zero, $P = 0.005$). Fuel consumption, however, showed no overall bias towards
positive or negative nonadditivity, but nonadditivity increased with time since
wetting (Fig. \ref{fig:mixture_obs_pred_consume}, slope not equal to zero,
$P < 0.0001$). Fuel consumption tended to show negative nonadditivity at higher
fuel moistures and positive nonadditivity following drying
(Fig.~\ref{fig:mixture_obs_pred_consume}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}

\subsection*{Moisture dynamics and flammability across taxonomic groups}

The taxa in this study displayed different patterns of both moisture absorption
and desorption, associated with different litter traits. In this sample of
species, those that have a greater absorptive capacity are also among the most
flammable in terms of spread rate when dry. Traits such as large leaves and
high SLA may affect both absorptive capacity and flammability with a
consequence that flammability rankings across species can change with time
since wetting. One shortcoming of past work in this system was that
flammability trials tested only dry fuels \citep{Magalhaes+Schwilk-2012}. We
therefore expected that rankings of species flammability might change under
higher moisture levels. For example, one presumption of ours prior to this
experiment was that the high flammability previously reported for \emph{Q.
  kelloggii} \citep{Magalhaes+Schwilk-2012} would disappear under higher fuel
moisture contents. This prediction was partly supported: \emph{Q. kelloggii}
fails to ignite until after 72 hours of drying in this study (Figs.
\ref{fig:spread_time} and \ref{fig:consume_time}). However, high moisture
content in this oak was unable to suppress ignition of mixtures to which it
contributed (Figs. \ref{fig:mixture_obs_pred_spread} and
\ref{fig:mixture_obs_pred_consume}).

The dry down curves we measured demonstrate that after a rain event, the litter
of different species will exhibit different moisture contents over time as a
consequence of both differential initial absorption and differences in drying
rate. This leads to flammability varying both by taxon and time since wetting.
In these mixed conifer forests, fine fuels can vary in fuel moisture throughout
the fire season. For example, 10-hour fuels in northern California varied from
less than 10\% to over 80\% depending on precipitation events in a study
following fuel moisture throughout the fire season
\citep{Estes+Knapp+etal-2012}. Early summer prescribed burns at our study sites
in Sequoia National Park show evidence of moisture limiting fire behavior at
12-14\% fuel moisture (Knapp et al 2005).

Our flammability trials demonstrate that both flame spread rate and fuel
consumption increased with time since wetting but that the rate of increase
varied across these taxa. Only the pines are able to ignite and support flame
spread immediately after wetting (Fig. \ref{fig:spread_time}). In contrast,
\emph{Quercus} exhibited a threshold response to drying with no ignition prior
to 72 hours of drying but once ignition is possible, it had faster spread rates
than the other taxa. Our two Cupressaceae, which have dense litter but have low
maximum water absorption, had low flammability across moisture levels. This
result is consistent with the low flammability found in an Australian
Cupressaceae that have small litter particles and dense packing
\citep{Scarff+Westoby-2006, Bowman+Haverkamp+etal-2018}. However, the densely
packed litter is not a consistent trait across the family because
\citet{Cornwell+Elvira+etal-2015} found relatively high litter flammability and
low litter density for some Asian Cupressaceae.

We examined one litter bed trait known to directly influence flammability,
litter bulk density, and one leaf trait, specific leaf area. Litter density is
largely driven by leaf size in this system and, in turn, controls the major
axis of flammability variation in dry litter \citep{Magalhaes+Schwilk-2012}.
Our results suggest that moisture dynamics are controlled by both traits
intrinsic to the leaf particles (specific leaf area) and an emergent property
of the litter bed (bulk density). Specific leaf area here is probably capturing
some joint effect of cuticle thickness that may influence water absorption and
total surface area of litter particles which increases water adsorption. The
oak has much higher maximum water absorption than do the other species,
followed by the finest-needled pine, \emph{P. lambertiana}. Our study examines
fresh litter, but we expect that decomposition will affect moisture dynamics.
For example, Van Wagner \citeyear{Van_Wagner-1969}, investigating fuel
weathering, found that pine needles had a variation in moisture content 3 to 6
times greater when they lacked their protective wax coating when compared to
intact needles.

\subsection*{Nonadditivity in litter mixtures}

We measured drying rates and flammability of four unique three-species litter
mixtures. These mixtures contained species representative of all dry-down
behavior types observed in single species trials. Mixtures that contained
\emph{Quercus} absorbed more water during wetting, and the one mixture without
any Cupressaceae had the most rapid drying rate (SI Figure 7 and SI Table 8).
Drying rates, however, did not behave as the average of the constituent
species: mixtures tended to be drier than expected early in the process and
wetter than expected later (Fig. \ref{fig:mixture_obs_pred_mc}). Past work has
shown that mixtures tend to pack less densely than do litter monocultures
\citep{Magalhaes+Schwilk-2012}. If packing density controls rate of drying,
then we would expect mixtures to dry more rapidly than predicted from
monoculture trials. Therefore, it is surprising that mixtures tended to have
lower maximum water content than predicted from the mean of the constituent
species, but higher water content than expected after long drying. This
discrepancy may be a consequence of the lightly packing litters having high
maximum water absorption and of drying being non-linear with time
(Fig.~\ref{fig:drydown}).

The overall ranking of flammability in litter mixtures showed some consistency
with the results obtained for the individual species, but there were
significant nonadditive effects. Mixtures containing \emph{Q. kelloggii} and
\emph{P. jeffreyi} tended to have fast spread rates. These results are likely
influenced by the lower bulk densities of litter mixtures containing those two
species. Most interestingly, there were patterns of nonadditivity in spread
rate and in fuel consumption for the four mixtures, indicating that positive
nonadditivity in flammability is not restricted to the completely dry litter
mixtures studied in previous work \citep{Magalhaes+Schwilk-2012,
  VanAltena+Logtestjin+etal-2012}. Nonadditivity in both spread rate and fuel
consumption became more positive with time since wetting and these patterns
were consistent across mixture identities. This is a slightly different result
than that reported by \citet{Blauw+Wensink+etal-2015}. Blauw et
al found that nonadditivity was greater in moist mixtures of heathland fuels
than in dry fuels. In their work, nonadditivity was generally negative but
cases of positive nonadditivity in mixtures was often a result of mixtures
meeting ignition thresholds not met by the constituent species in monoculture.
To limit the number of models fit, we limited our investigation to two
flammability parameters that represent two major axes of flammability
variation. Therefore, our study does not fully capture ignition effects
although we do have some results consistent with theirs: mixtures that included
\emph{Q. kelloggii} could ignite at times since wetting at which \emph{Quercus}
in monoculture could not.

Overall, we report a consistent pattern of increasingly positive nonadditivity
in flammability with litter drying across two largely orthogonal measures of
flammability \citep{Schwilk-2015, Pausas+Keeley+etal-2017,
  Prior+Murphy+etal-2018}. Some mixtures, however did exhibit negative
nonadditivity. For example, at 72 hours following saturation, fuel consumption
across mixtures showed neutral, positive, or negative nonadditivity (Fig.
\ref{fig:mixture_obs_pred_consume}). Previous studies have demonstrated the
potential for the direction of nonadditivity to shift with fuel moisture
\citep{Blauw+Wensink+etal-2015} or with altering mixtures of leaf vs twig
litter \citep{Zhao+vanLogtestijn+etal-2019}. Our litter represented naturally
dropped litter and the ratio of leaves to twigs in such material can vary both
by species but also by wind and storm events the previous season. With only
four distinct litter combinations, it is impossible to attribute cause to these
variations within the overall statistical trend, but litter packing alone is
complex and needles and twigs can combine such that the addition of a species
to a mixture can create a denser litter bed or less dense litter bed than is
expected \citep{Zhao+vanLogtestijn+etal-2019}.

% (eg \emph{Qu)(eg \emph{Abies} branchlets with needles fitting between loosely
% packed \emph{Pinus} needles) or less dense

We draw two main conclusions from this work. First, maximum water absorption is
highest in those species with the most flammability litter under dry conditions.
As a consequence, under higher fuel moisture, the flammability ranking of
these taxa can change compared to the flammability of dry fuels. Second, litter
flammability is prone to nonadditive effects which become more positive with
drying. The relationships we described here can form a working hypothesis and
baseline expectation to adjust fire behavior predictions in real communities:
in mixtures with fuel moisture content below 50\%, observed flame spread rates
approached double the predicted value based on monoculture trials (Fig.
\ref{fig:mixture_obs_pred_spread}). In completely dry litter from this same
system, past work demonstrates observed spread rates approached three times the
expectation \citep{Magalhaes+Schwilk-2012}. These numbers provide a starting
point for parameterizing fire models with mixed fuels. In this temperate mixed
conifer forest, these effects are likely to include positive feedbacks whereby
the shade intolerant species such as the long-needled pines and the
large-leaved oak, are favored by the increased fire to which their fuels
contribute \citep{Schwilk+Caprio-2011} and similar results are reported for
southeastern US forests where species composition shifts have led to decreases
in fire activity \citep{Nowacki+Abrams-2008}. Vegetation-fire feedbacks are
important at a wide range of scales from stand-level species composition shifts
to the earth system \citep{Harris+Remenyi+etal-2016,
  Archibald+Lehmann+etal-2018}. The patterns we demonstrate here show the
potential importance of species traits and interactions at local and stand
scales where management and fire operations meet species' biology.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgments}

The authors thank the National Park Service and the employees of Sequoia and
Kings Canyon National Parks. This work was carried out with permission from NPS
under permit \# SEKI-2014-SCI-0040. The authors thank Russell Lackey who
contributed essential assistance with the field work and Gabrielle Plata and
Leticia Plata who assisted with the burning trials. We thank the Lubbock Fire
Department for their ongoing help and cooperation.

This research received no external funding.

\section*{Author contributions}

Conceptualization, RMQM and DWS; methodology, RMQM and DWS; experiments, RMQM;
coding and data analysis, DWS and RMQM; writing DWS and RMQM; writing--review
and editing, DWS

\section*{Data and code accessibility}
All data and code for this project is at
\url{https://github.com/schwilklab/trait-flam}. DOI to be provided. For review,
use ``joe'' branch which contains this manuscript and code needed to create the
figures and tables.

\newpage
\bibliography{Magalhaes_Schwilk_litter_flam_moisture.bib}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \conflictsofinterest{The authors declare no conflict of interest. The funders
% had no role in the design of the study; in the collection, analyses, or
% interpretation of data; in the writing of the manuscript, or in the decision
% to publish the results.}

% References, variant B: external bibliography
% =====================================

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}